Self Regulation Ontology Retest Analyses

### 2018-03-21 code/SRO_Retest_Analyses.Rmd
refactoring

```{r, message=FALSE, warning=FALSE, include=FALSE}
source('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/code/SRO_Retest_Analyses_Helper_Functions.R')

@@

```{r}
p1_t_legend <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='black')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'bottom',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30), 
        legend.text = element_text(size=20), 
        legend.key.width = unit(0.75, "inches"), 
        legend.title = element_text(size=28),
        legend.spacing.x = unit(0.5, "inches")) + 
  guides(size = guide_legend(override.aes = list(size=c(9,18,28))),
         shape = guide_legend(override.aes = list(size=18)))+
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3), name="Type")+
  scale_size_continuous(name = "Sample Size")+
  geom_vline(xintercept = 0, color = "red", size = 1)

p1_t <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#00BFC4')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

p2_t <- lit_review %>%
  filter(task == 'survey') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#F8766D')+
  theme_bw()+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

@@

```{r warning=FALSE, message=FALSE}
t1_2_difference %>%
  ggplot(aes(difference, alpha=dv))+
  geom_histogram(position='identity')+
  theme_bw()+
  theme(legend.position = 'none')
```

@@

```{r}
t1_2_difference %>%
  ggplot()+
  geom_smooth(aes(as.numeric(days_btw), abs(difference), group=factor(dv)), method='lm', se=FALSE)+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme_bw()+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Absolute Scaled difference score')

@@

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), abs(difference), 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme_bw()+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))
```

@@

```{r}
sig_days_effect %>%
  left_join(rel_df, by='dv') %>%
  ggplot(aes(estimate, icc))+
  geom_point()+
  theme_bw()
```

@@

```{r}
var_boot_df = boot_df %>%
  group_by(dv) %>%
  summarise(icc_median = quantile(icc, probs = 0.5),
            icc_2.5 = quantile(icc, probs = 0.025),
            icc_97.5 = quantile(icc, probs = 0.975)
            )

boot_lit_rel = function(df, samples = 100, rel_col = 'retest_reliability'){

  boot_once = function(df){
    #get number of rows in df
    data_amount = nrow(df)
    #sample as many reliability estimates as there are rows in the df
    boot_sample = sample(c(1:data_amount), size=data_amount, replace=TRUE)
    boot_sample = df[boot_sample,]
    #calculate median reliability estimate for the sampled estimates
    return(median(boot_sample[,rel_col])) 
  }

  out = replicate(samples, boot_once(df))

  return(data.frame(median_lit_rel=out))
}

var_lit_review = lit_review %>%
  group_by(dv)%>%
  do(boot_lit_rel(.))
  summarise(median_reliability=median(retest_reliability),
            task = unique(task)) %>%
  mutate(dv = as.character(dv))

var_boot_df = boot_df %>%
  group_by(dv) %>%
  summarise(median_icc = quantile(icc, probs = 0.5))

var_lit_review = lit_review %>%
  # filter(sample_size > 100) %>%
  group_by(dv)%>%
  summarise(median_reliability=median(retest_reliability),
            task = unique(task),
            mean_sample = mean(sample_size)) %>%
  mutate(dv = as.character(dv))

rel_comp = var_lit_review %>%
  left_join(var_boot_df, by = 'dv')

rel_comp
```

Correlation between our median estimates from bootstrapped samples and the literature review for task variables
```{r}
with(rel_comp %>% filter(task == "task"), cor(median_icc, median_reliability))
```

```{r}
with(rel_comp %>% filter(task == "survey"), cor(median_icc, median_reliability))
```

@@

```{r}
rel_comp %>%
  ggplot(aes(median_reliability, median_icc, color=task))+
  geom_point()+
  geom_abline(slope=1, intercept=0)+
  theme_bw()+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  theme(legend.title = element_blank())+
  xlab('Median reliability from literature')+
  ylab('Median ICC from bootstrapped samples')
```

```{r}
lit_review %>%
  ggplot(aes(sample_size, retest_reliability, color=task))+
  geom_point()
summary(lmer(retest_reliability~sample_size+task+(1|dv), data=lit_review))

#sd by average sample size
lit_review %>%
  group_by(dv) %>%
  summarise(sd_rel = sd(retest_reliability),
            avg_sample_size = mean(sample_size)) %>%
  ggplot(aes(avg_sample_size, sd_rel))+
  geom_point()+
  xlim(0,200)
```

x = deviance between our data and literature
y = average sample size

```{r}
rel_comp %>%
  mutate(diff_sq = (median_icc - median_reliability)^2) %>%
  ggplot(aes(diff_sq, mean_sample, color=task))+
  geom_point()+
  geom_smooth()
```

```{r}
lit_review %>%
  ggplot(aes(sample_size))+
  geom_histogram()+
  theme_bw()
```

@@

```{r}
process_boot_df = function(df){
  df = df %>%
  drop_na() %>%
  mutate(dv = as.character(dv),
         icc = as.numeric(as.character(icc)),
         spearman = as.numeric(as.character(spearman)),
         pearson = as.numeric(as.character(pearson)),
         eta_sq = as.numeric(as.character(eta_sq)),
         sem = as.numeric(as.character(sem)),
         partial_eta_sq = as.numeric(as.character(partial_eta_sq)),
         omega_sq = as.numeric(as.character(omega_sq)),
         var_subs = as.numeric(as.character(var_subs)),
         var_ind = as.numeric(as.character(var_ind)),
         var_resid = as.numeric(as.character(var_resid)),
         F_time = as.numeric(as.character(F_time)),
         p_time = as.numeric(as.character(p_time)),
         df_time = as.numeric(as.character(df_time)),
         df_resid = as.numeric(as.character(df_resid)))
  return(df)} 
```

```{r message=FALSE, warning=FALSE}
boot_df <- read.csv(gzfile(paste0(retest_data_path,'bootstrap_merged.csv.gz')))

rm(refit_boot_df)

>>>>

```{r, message=FALSE, warning=FALSE, include=FALSE}
source('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/code/SRO_Retest_Analyses_Helper_Functions.R')
theme_set(theme_bw())

@@

```{r}
p1_t_legend <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='black')+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'bottom',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30), 
        legend.text = element_text(size=20), 
        legend.key.width = unit(0.75, "inches"), 
        legend.title = element_text(size=28),
        legend.spacing.x = unit(0.5, "inches")) + 
  guides(size = guide_legend(override.aes = list(size=c(9,18,28))),
         shape = guide_legend(override.aes = list(size=18)))+
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3), name="Type")+
  scale_size_continuous(name = "Sample Size")+
  geom_vline(xintercept = 0, color = "red", size = 1)

p1_t <- lit_review %>%
  filter(task == 'task') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#00BFC4')+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

p2_t <- lit_review %>%
  filter(task == 'survey') %>%
ggplot(aes(y = factor(task_group, levels=rev(unique(task_group))), x = retest_reliability)) + 
  geom_point(aes(size=sample_size, shape = type), color='#F8766D')+
  theme(axis.text.y = element_text(size=43),
        legend.position = 'none',
        axis.text.x = element_text(size=23),
        axis.title.x = element_text(size=30)) + 
  xlab("Reliability")+
  ylab("")+
  scale_x_continuous(limits = c(-0.25,1), breaks=c(-0.25, 0, 0.25, 0.5, 0.75, 1))+
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10))+
  scale_shape_manual(breaks = sort(lit_review$type), values = c(15, 16, 17, 3))+
  scale_size_continuous(range = c(5, 35))+
  geom_vline(xintercept = 0, color = "red", size = 1)

@@

```{r warning=FALSE, message=FALSE}
t1_2_difference %>%
  ggplot(aes(difference, alpha=dv))+
  geom_histogram(position='identity')+
  theme(legend.position = 'none')
```

@@

```{r}
t1_2_difference %>%
  ggplot()+
  geom_smooth(aes(as.numeric(days_btw), abs(difference), group=factor(dv)), method='lm', se=FALSE)+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Absolute Scaled difference score')

@@

```{r}
t1_2_difference %>%
  mutate(sig_days_effect = ifelse(dv %in% sig_days_effect$dv, 1, 0))%>%
  arrange(-sig_days_effect) %>%
  ggplot()+
  stat_smooth (aes(as.numeric(days_btw), abs(difference), 
                  group=factor(dv, levels=unique(dv[order(sig_days_effect)]), ordered=TRUE), 
                  color=factor(sig_days_effect, levels = c(1,0))),geom="line", alpha=0.5, method='lm')+
  geom_smooth(aes(as.numeric(days_btw), abs(difference)), method='lm', color = "black", se=FALSE)+
  theme(legend.title = element_blank())+
  xlab('Days between completion')+
  ylab('Scaled difference score')+
  scale_color_discrete(breaks=c(0,1),
                       labels=c("NS", "Sig"))
```

@@

```{r}
sig_days_effect %>%
  left_join(rel_df, by='dv') %>%
  ggplot(aes(estimate, icc))+
  geom_point()
```

`Plot differences for the measures with differences as examples`

@@

Read in and process bootstrapped results.

```{r}
process_boot_df = function(df){
  df = df %>%
  drop_na() %>%
  mutate(dv = as.character(dv),
         icc = as.numeric(as.character(icc)),
         spearman = as.numeric(as.character(spearman)),
         pearson = as.numeric(as.character(pearson)),
         eta_sq = as.numeric(as.character(eta_sq)),
         sem = as.numeric(as.character(sem)),
         partial_eta_sq = as.numeric(as.character(partial_eta_sq)),
         omega_sq = as.numeric(as.character(omega_sq)),
         var_subs = as.numeric(as.character(var_subs)),
         var_ind = as.numeric(as.character(var_ind)),
         var_resid = as.numeric(as.character(var_resid)),
         F_time = as.numeric(as.character(F_time)),
         p_time = as.numeric(as.character(p_time)),
         df_time = as.numeric(as.character(df_time)),
         df_resid = as.numeric(as.character(df_resid)))
  return(df)}

boot_df <- read.csv(gzfile(paste0(retest_data_path,'bootstrap_merged.csv.gz')))

rm(refit_boot_df)
```

Summarize bootstrapped results and merge to lit review data

```{r}
var_boot_df = boot_df %>%
  group_by(dv) %>%
  summarise(mean_icc = mean(icc))

rel_comp = lit_review %>%
  left_join(var_boot_df, by = 'dv')
```

Here's what our data looks like: (583 data points for 171 measures)

```{r}
rel_comp
```

### Lit review results

Distribution of reliabilities, sample sizes and delays

```{r}
rel_comp %>% 
  select(dv, task, retest_reliability, sample_size, days) %>%
  filter(days < 3600) %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value, fill=task))+
  geom_density(alpha=0.5, position='identity')+
  facet_wrap(~key, scales='free')+
  theme(legend.title = element_blank())
```

What predicts retest reliability in the literature?
Task, sample size, days

```{r}
mod1 = lmer(retest_reliability ~ task + (1|dv), rel_comp)
mod2 = lmer(retest_reliability ~ task + sample_size + (1|dv), rel_comp)

anova(mod1, mod2)
```

```{r}
mod3 = lmer(retest_reliability ~ task + sample_size + days+ (1|dv), rel_comp)

anova(mod2, mod3)
```

```{r}
summary(mod2)
```

```{r}
rel_comp %>% 
  ggplot(aes(sample_size, retest_reliability, color=task))+
  geom_smooth(method='lm')+
  geom_point(alpha = 0.2)+
  theme(legend.title = element_blank())+
  xlab("Sample Size")+
  ylab("Retest reliability")
```

Compare effect sizes of effect of task on reliability estimates in literature vs our results (just looking at the variables you find in the literature)

### Conceptual replications 
- Larger reliability estimates for surveys
- Larger variance in reliability estimates for tasks

### Sample size effects/publication bias

Deviance between our mean estimate and the reliability from the literature decreases exponentially with sample size. The smaller the sample size in the literature the more the reliability estimate differs from our results.

```{r}
rel_comp = rel_comp %>%
  mutate(diff_sq = abs(mean_icc - retest_reliability)^2)

rel_comp %>%
  ggplot(aes(sample_size,diff_sq, color=task))+
  # geom_point()+
  #exponential decay (linearized by taking logs of both sides)
  geom_smooth(method="glm",se=FALSE, method.args=list(family=gaussian(link="log")))+
  theme_bw()
```

```{r}
lit_review %>%
  ggplot(aes(sample_size, retest_reliability, color=task))+
  geom_smooth(method='lm')

mod1 = lmer(retest_reliability~sample_size+(1|dv), data=lit_review)

```{r}
#sd by average sample size
lit_review %>%
  group_by(dv) %>%
  summarise(sd_rel = sd(retest_reliability),
            avg_sample_size = mean(sample_size)) %>%
  ggplot(aes(avg_sample_size, sd_rel))+
  geom_point()+
  xlim(0,200)
```

### Direct relation to our results

Correlation between our mean estimates from bootstrapped samples and the literature review for task variables

```{r}
with(rel_comp %>% filter(task == "task"), cor(mean_icc, mean_reliability))
```

```{r}
with(rel_comp %>% filter(task == "survey"), cor(mean_icc, mean_reliability))
```

Model comparisons building model to predict the reliabilities in the literature from our results.

```{r}
mod1 = lmer(retest_reliability ~ mean_icc + (1|dv), data=rel_comp)
mod2 = lmer(retest_reliability ~ mean_icc + (mean_icc|dv), data=rel_comp)

anova(mod1, mod2)
```

```{r}
mod3 = lmer(retest_reliability ~ mean_icc + scale(sample_size) + (mean_icc|dv), data=rel_comp)

anova(mod2, mod3)
```

```{r}
mod4 = lmer(retest_reliability ~ mean_icc + scale(days) + scale(sample_size) + (mean_icc|dv), data=rel_comp)

anova(mod3, mod4)
```

Selected model: fixed effect for mean icc and sample size of study, random slope with respect to mean icc and random intercept for each measure

What does it say?

```{r}
summary(mod3)
```

How well does the model fit? (squared correlation between predictions of model and data)

```{r}
cor(predict(mod3), rel_comp$retest_reliability)^2
```

Same result if you do flat model without random slopes for dv's

```{r}
summary(lm(retest_reliability ~ mean_icc+scale(sample_size), data=rel_comp))
```

Visualize hierarchical model:
How would I show random intercepts/slopes? Does the fact that I can't think of how means that there is something wrong with the model?

```{r}
rel_comp %>%
  ggplot()+  
  geom_boxplot(aes(mean_icc, retest_reliability, group=dv))+
  geom_smooth(aes(mean_icc, retest_reliability), method= 'lm')+
  theme_bw()+
  theme(legend.position = 'none')+
  geom_abline(slope=1, intercept=0)+
  # geom_abline(slope=0, intercept=0.3, linetype='dashed')+
  xlim(-0.2, 1)+
  ylim(-0.2, 1)+
  facet_grid(task~.)+
  xlab('Mean bootstrapped ICC')+
  ylab('Reliability estimate from literature')

@@

@@

```{r message=FALSE, warning=FALSE}
